Supplementary Information: Quantifying the pathways to life using assembly spaces

### 2021-05-14 papers/assembly/I/main.tex
[papers] Tag necessary changes related to the definition of a pathway

Up to this point, all of the defintions have been fairly standard, at least in terms of concepts.
We now start constructing apparently new ideas that will allow us to formalize the concepts of an assembly space and the assembly index.
The first of these is the idea of a pathway.
Suppose you begin with a given object $x$ and wish to fleshout all of the ways in which a second object $y$ might be constructed.
The simplest first step is to construct a sequence of transitions, the first of which takes $x$ as part of its input, the last yields $y$ as part of it its output, and every transition in between depends at least partially on the output of the previous.
Such a sequence of transitions is what we refer to as a \textit{pathway}.

\begin{definition}\label{def:pathway}
    If $P$ is a Petri net, a \textbf{pathway} $\gamma = t_n \ldots t_1$ in $P$ of length $n \ge 1$ is a sequence of transitions $t_i \in T$ such that $\tau(t_i) \ne 0$, $\sigma(t_i) \ne 0$, $m_{i} + \tau(t_i) = n_{i+1} + \sigma(t_{i+1})$ with $\sigma(t_{i+1}) \ne m_{i}$ and $\tau(t_{i+1}) \ne n_{i+1}$. We write $t \in \gamma$ to denote that the transition $t$ appears in the pathway $\gamma$.
\end{definition}

@@

\begin{definition}\label{def:pathway-connected}
    Let $P$ be a Petri net and $x,y \in S(P)$.
    We say that $x$ is \textbf{pathway connected to} $y$ if and only if there exists a pathway $\gamma = t_n \ldots t_1$ such that $\sigma(t_1) = x + m$ and $\tau(t_n) = y + n$ for some $m, n \in \N[S(P)]$.
    The set of all pathways in $P$ from $x$ to $y$ is denoted $\Gamma_x^y$.
    We denote the set of all pathways in $P$ starting from $x$ as $\Gamma_x = \bigcup_{y \in S(P)} \Gamma_x^y$, and all pathways terminating at $y$ as $\Gamma^y = \bigcup_{x \in S(P)} \Gamma_x^y$.
\end{definition}

In our ligation example, you might suppose that you begin with $A$ and would like to ultimately yield $AAAB$.
There are in fact two pathways which accomplish this, $\gamma\alpha$ and $\gamma\beta$, which we read as \textit{first fire $\alpha$ ($\beta$, respectively) and then fire $\gamma$}:
\begin{center}
    \scalebox{0.6}{
        \begin{tikzpicture}
            \begin{pgfonlayer}{nodelayer}
                \node [style=site] (A) at (0, 0.8) {$A$};
                \node [style=site] (B) at (0, -0.8) {$B$};
                \node [style=site] (C) at (6, -1.6) {$C$};
                \node [style=site] (AA) at (4, 0.8) {$AA$};
                \node [style=site] (AB) at (4, -0.8) {$AB$};
                \node [style=site] (AAAB) at (8, 0) {$AAAB$};
                \node [style=transition, fill=green!20] (tau1) at (2, 0.8) {$\alpha$};
                \node [style=transition] (tau2) at (2, -0.8) {$\beta$};
                \node [style=transition, fill=green!20] (tau3) at (6, 0) {$\gamma$};
                \node [style=transition] (tau4) at (10, 0) {$\delta$};
                \node [scale=3] at (5, -3.5) {$\gamma\alpha$};

@@

As with subnets, pathways interact well with Petri net morphisms.
This brings us to our first lemma:
\begin{lemma}\label{lem:image-of-pathways}
    Let $f: P \rightarrow Q$ be a Petri net morphism and let $x, y \in S(P)$ be pathway connected in $P$.
    Then $f(x)$ and $f(y)$ are pathway connected in $f(P)$.
\end{lemma}
\begin{proof}
    Let $\gamma = t_n \ldots t_1$ be a pathway from $x$ to $y$.
    Given that $\sigma_P(t_i) \ne 0$ and $\tau_P(t_i) \ne 0$, we know that $\sigma_{f(P)} \ne 0$ and $\tau_{f(P)} \ne 0$.
    Since $\sigma_P(t_1) = x + m$, $\sigma_{f(P)}(f(x)) = f(x) + m'$ for some $m' \in \N[S(f(P))]$; similarly, $\tau_{f(P)}(f(y)) = f(y) + n'$ for some $n' \in \N[S(f(P))]$.

>>>>

Up to this point, all of the defintions have been fairly standard, at least in terms of concepts.
We now start constructing apparently new ideas that will allow us to formalize the concepts of an assembly space and the assembly index.
The first of these is the idea of a pathway.
Suppose you begin with a given object $x$ and wish to fleshout all of the ways in which a second object $y$ might be constructed.
The simplest first step is to construct a sequence of transitions, the first of which takes $x$ as part of its input, the last yields $y$ as part of it its output, and every transition in between depends at least partially on the output of the previous.
Such a sequence of transitions forms the core of what we refer to as a \textit{pathway}.

{
    \color{red}
    \begin{definition}\label{def:pathway}
        If $P$ is a Petri net, a \textbf{pathway} $\gamma = t_n \ldots t_1$ in $P$ of length $n \ge 1$ is a sequence of transitions $t_i \in T$ such that $\tau(t_i) \ne 0$, $\sigma(t_i) \ne 0$, $m_{i} + \tau(t_i) = n_{i+1} + \sigma(t_{i+1})$ with $\sigma(t_{i+1}) \ne m_{i}$ and $\tau(t_{i+1}) \ne n_{i+1}$. We write $t \in \gamma$ to denote that the transition $t$ appears in the pathway $\gamma$.
    \end{definition}
}

@@

\begin{definition}\label{def:pathway-connected}
    \TODO{Update this to reflect changes in the definition of a pathway.}
    Let $P$ be a Petri net and $x,y \in S(P)$.
    We say that $x$ is \textbf{pathway connected to} $y$ if and only if there exists a pathway $\gamma = t_n \ldots t_1$ such that $\sigma(t_1) = x + m$ and $\tau(t_n) = y + n$ for some $m, n \in \N[S(P)]$.
    The set of all pathways in $P$ from $x$ to $y$ is denoted $\Gamma_x^y$.
    We denote the set of all pathways in $P$ starting from $x$ as $\Gamma_x = \bigcup_{y \in S(P)} \Gamma_x^y$, and all pathways terminating at $y$ as $\Gamma^y = \bigcup_{x \in S(P)} \Gamma_x^y$.
\end{definition}

\TODO{Update this example to reflect changes in the definition of a pathway.}
In our ligation example, you might suppose that you begin with $A$ and would like to ultimately yield $AAAB$.
There are in fact two pathways which accomplish this, $\gamma\alpha$ and $\gamma\beta$, which we read as \textit{first fire $\alpha$ ($\beta$, respectively) and then fire $\gamma$}:
\begin{center}
    \scalebox{0.6}{
        \begin{tikzpicture}
            \begin{pgfonlayer}{nodelayer}
                \node [style=site] (A) at (0, 0.8) {$A$};
                \node [style=site] (B) at (0, -0.8) {$B$};
                \node [style=site] (C) at (6, -1.6) {$C$};
                \node [style=site] (AA) at (4, 0.8) {$AA$};
                \node [style=site] (AB) at (4, -0.8) {$AB$};
                \node [style=site] (AAAB) at (8, 0) {$AAAB$};
                \node [style=transition, fill=green!20] (tau1) at (2, 0.8) {$\alpha$};
                \node [style=transition] (tau2) at (2, -0.8) {$\beta$};
                \node [style=transition, fill=green!20] (tau3) at (6, 0) {$\gamma$};
                \node [style=transition] (tau4) at (10, 0) {$\delta$};
                \node [scale=3] at (5, -3.5) {$\gamma\alpha$};

@@

As with subnets, pathways interact well with Petri net morphisms.
This brings us to our first lemma:
\begin{lemma}\label{lem:image-of-pathways}
    \TODO{Update this proof to reflect changes to the definition of a pathway.}
    Let $f: P \rightarrow Q$ be a Petri net morphism and let $x, y \in S(P)$ be pathway connected in $P$.
    Then $f(x)$ and $f(y)$ are pathway connected in $f(P)$.
\end{lemma}
\begin{proof}
    Let $\gamma = t_n \ldots t_1$ be a pathway from $x$ to $y$.
    Given that $\sigma_P(t_i) \ne 0$ and $\tau_P(t_i) \ne 0$, we know that $\sigma_{f(P)} \ne 0$ and $\tau_{f(P)} \ne 0$.
    Since $\sigma_P(t_1) = x + m$, $\sigma_{f(P)}(f(x)) = f(x) + m'$ for some $m' \in \N[S(f(P))]$; similarly, $\tau_{f(P)}(f(y)) = f(y) + n'$ for some $n' \in \N[S(f(P))]$.
